Refinements for open automata

### 2023-04-25 2021-Refinement/paper/refinementarticleV2.tex
refinements

In the following definition we add a parameter $H$ which is the set of holes tracked by the refinement relation and adapt the definition by ignoring actions of the holes that are not in $H$.
\begin{definition}[Open automata refinement]\label{Def:OA-Refinement}
For two open automata\\ \(A_1 \defobject \OA{S_1}{s_{01}}{J_1}{V_1}{\sigma_{01}}{T_1}\) and \(A_2 \defobject \OA{S_2}{s_{02}}{J_2}{V_2}{\sigma_{02}}{T_2}\), \(A_1\) is a refinement of \(A_2\) tracking holes \(H\), noted \(\wrel{A_1}{A_2}{H}\), with \(H \subseteq J_1 \cap J_2\), if there is a relation on configurations $R: \mpar{S_1 \times S_2} \to \rformulas[V_1 \uplus V_2]$ such that:
\item[(1)] \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\)
\item[(2)] \(\forall \mpar{s_1, s_2} \in S_1 \times S_2,\)
\begin{multline*}
	\everymath{\displaystyle}\begin{array}{l}
		\bigsymb{\forall} \OTx{1}{}{1}{1} \in \fOT{s_1}, \bigsymb{\exists} \mpar{\OTx{2}{x}{2x}{2x} \in \fOT{s_2}}^{x \in X}, \\[12pt]
		\quad \mpar{\forall x \in X, J'_{2x} \cap H = J'_1 \cap H} \nwedge\\[1pt]
		 V_1 \uplus V_2 \uplus \fvars{t_1} \vdash\\\hspace{5em} \mpar{R\mpar{s_1, s_2} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\begin{array}{l}
			\alpha_1 = \alpha_{2x} \wedge \bigwedge_\subbox{j \in J'_{2x} \cap H} \beta_{1j} = \beta_{2xj}   \nwedge\\
			 g_{2x} \wedge R\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}}
		\end{array}}} 
	\end{array} 
\end{multline*}
\item[(3)] Deadlock reduction:
\[
\forall \mpar{s_1, s_2} \in S_1 \times S_2, V_1 \uplus V_2  \vdash R\mpar{s_1, s_2} \implies( \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} ) \lor¢ \lnot( \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2}  )
\]
\end{definition}
Note that every action of the holes outside \(H\) is unconstrained according to the refinement relation.

\begin{example} It is easy to see the  open automaton of Fig.~\ref{Fig:RefineOA} is a refinement the open automaton of Fig.~\ref{Fig:SpecOA} according to this definition.
\end{example}

@@

\begin{lemma}[Tracked holes] By construction, if an automaton is the refinement of another one, it is also a refinement by tracking less holes.
\[\wrel{A_1}{A_2}{H} \land H'\subseteq H \implies \wrel{A_1}{A_2}{H'}\]
\end{lemma}

@@

We now  state the properties of our refinement relations, proofs of the properties can be found in the appendices. We express these properties in terms of open automata refinement because hole-equal simulation is a particular case of Definition~\ref{Def:OA-Refinement} when  $J_1=J_2=H$, and thus all the properties shown here are trivially true for hole-equal simulation.
The first crucial property of  refinement  is that it is  a  preorder on the set of open automata. This property enables stepwise refinement.

@@

\begin{example}  Consider again the \texttt{prod-cons}  \texttt{Simprod-cons}  automata given in examples above.  Since we have \texttt{prod-cons} $\leq_{\{\texttt{P,Q}\}}$ \texttt{simprod-cons},  hence according to Theorem \ref{thm:ContextRefinement}, we get 
\texttt{prod-cons[P/producer]} $\leq_{\{\texttt{Q}\}}$  \texttt{Simprod-cons[P/producer]}. This can be verified by comparing the behaviour of the automaton of the figure Fig.~\ref{Fig:ComposeOA} and the automaton resulting  from the composition of the \texttt{Simprod} and  \texttt{producer} automata.\qed
\end{example}

>>>>

In the following definition we add a parameter $H$ which is the set of holes tracked by the refinement relation and adapt the definition by ignoring actions of the holes that are not in $H$.
\begin{definition}[Hole-tracking simulation]\label{Def:OA-Refinement}
For two open automata\\ \(A_1 \defobject \OA{S_1}{s_{01}}{J_1}{V_1}{\sigma_{01}}{T_1}\) and \(A_2 \defobject \OA{S_2}{s_{02}}{J_2}{V_2}{\sigma_{02}}{T_2}\), \(A_1\) is a refinement of \(A_2\) tracking holes \(H\), noted \(\wrel{A_1}{A_2}{H}\), with \(H \subseteq J_1 \cap J_2\), if there is a relation on configurations $R: \mpar{S_1 \times S_2} \to \rformulas[V_1 \uplus V_2]$ such that:
\item[(1)] \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\)
\item[(2)] \(\forall \mpar{s_1, s_2} \in S_1 \times S_2,\)
\begin{multline*}
	\everymath{\displaystyle}\begin{array}{l}
		\bigsymb{\forall} \OTx{1}{}{1}{1} \in \fOT{s_1}, \bigsymb{\exists} \mpar{\OTx{2}{x}{2x}{2x} \in \fOT{s_2}}^{x \in X}, \\[12pt]
		\quad \mpar{\forall x \in X, J'_{2x} \cap H = J'_1 \cap H} \nwedge\\[1pt]
		 V_1 \uplus V_2 \uplus \fvars{t_1} \vdash\\\hspace{5em} \mpar{R\mpar{s_1, s_2} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\begin{array}{l}
			\alpha_1 = \alpha_{2x} \wedge \bigwedge_\subbox{j \in J'_{2x} \cap H} \beta_{1j} = \beta_{2xj}   \nwedge\\
			 g_{2x} \wedge R\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}}
		\end{array}}} 
	\end{array} 
\end{multline*}
\item[(3)] Deadlock reduction:
\[
\forall \mpar{s_1, s_2} \in S_1 \times S_2, V_1 \uplus V_2  \vdash R\mpar{s_1, s_2} \implies( \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} ) \lor¢ \lnot( \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2}  )
\]
\end{definition}
Note that every action of the holes outside \(H\) is unconstrained according to the refinement relation.

%\begin{example} It is easy to see the  open automaton of Fig.~\ref{Fig:RefineOA} is a refinement the open automaton of Fig.~\ref{Fig:SpecOA} according to this definition.
%\end{example}

\begin{property}[Relating simulations]\label{lem-sim-sim}
Hole-equal simulation is a particular case of hole-tracking refinement when  $J_1=J_2=H$.
\end{property}
\TODO{potential ex 4: check if the example with a hole filled is a simulation of the example with holes? at least state \texttt{prod-cons} $\leq_{\{\texttt{P,Q}\}}$ \texttt{simprod-cons}}

@@

\begin{property}[Tracked holes] By construction, if an automaton is the refinement of another one, it is also a refinement by tracking less holes.
\[\wrel{A_1}{A_2}{H} \land H'\subseteq H \implies \wrel{A_1}{A_2}{H'}\]
\end{property}

@@

We now  state the properties of our refinement relations, proofs of the properties can be found in the appendices. We express these properties in terms of hole-tracking simulation because, thanks to Property~\ref{lem-sim-sim}   all the properties  of hole-tracking simulation are also valid  for hole-equal simulation
The first crucial property of  refinement  is that it is  a  preorder on the set of open automata. This property enables stepwise refinement.

@@

\begin{example}  Consider again the \texttt{prod-cons} and \texttt{Simprod-cons}  automata given in the examples above.  Since we have \texttt{prod-cons} $\leq_{\{\texttt{P,Q}\}}$ \texttt{simprod-cons},  hence according to Theorem \ref{thm:ContextRefinement}, we get 
\texttt{prod-cons[P/producer]} $\leq_{\{\texttt{Q}\}}$  \texttt{Simprod-cons[P/producer]}. This can be verified by comparing the behaviour of the automaton of the figure Fig.~\ref{Fig:ComposeOA} and the automaton resulting  from the composition of the \texttt{Simprod} and  \texttt{producer} automata.\qed
\end{example}
